### 2020-12-20 A-BasicCommom/ACL2021/acl_summary.tex
1. Summarization for Document.

%
% File acl2020.tex
%
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{A Transformer-based Pretrained Language Model for Automatic Document Summarization}

\author{Dezhou Shen \\
  Department of Computer Science\\
  Tsinghua University\\
  Beijing, CN 100084\\
  \texttt{sdz15@mails.tsinghua.edu.cn} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
 Document summarization is an essential task in the information retrieval.
 User intention for specific documents is hard to satisfy for the information overload,
 thus there are rarely golden standards for matching a certain document.
 I propose a multi-grained summarization model for documents that uses attention on paragraphs,
 supervised learning on the sentence importance, and sequence-to-sequence learning on characters and words.
 Tested on the 2020 Chinese Law Summarization Dataset, the proposed model achieved
 better performance than other modern summarization models.

\end{abstract}

\section{Introduction}

Document matching is an essential task in daily work and study.
Because the matching algorithm is not capable of choosing important words from the massive documents,
to use the content of a document as the matching criteria often fails in returning unintended documents.
Document summarization condenses long documents into short paragraphs by retaining core information.
Thus, a much wiser way for the document matching is using the summarization instead of the content.

\section{Related Work}

\citet{bhargava2020deep} use Convolutional neural network and Long-Short Term Memory network in the paraphrase detection of
documents and summaries, then using multi-grained Convolutional neural networks as the sentence extraction model,
as a result, the F1 scores on the English and Malayalam summarization corpus are 31.57\% and 53.57\%.

\citet{erera2019summarization} consider the summarization problem as a query problem, use query
saliency, entities coverage, diversity, text coverage and sentence length as the objectives.

\citet{altmami2020automatic} concluded the special pattern for the summarization of scientific articles,
and found that abstract generation-based and citation-based are two notice-worth solutions.

\citet{polsley2016casesummarizer} used a combined word frequency and domain knowledge summary method for the legal text abstraction,
and achieved an average of 10.9\% ROUGE score between automatically-generated summaries, and the expert-generated summaries.

\citet{iqbal2020survey} showed that the deep learning approaches have achieved many satisfactory results in several domains,
Word embeddings, such as Word2Vec, Glove, and FastText are widely used in the Natural Language Processing.
Variational Auto-Encoders and Generative Adversarial Networks have been used for the text generations.

\citet{litvak2019easy} studied the evaluation system for text summarization,
and showed that ROUGE, MeMoG for quality evaluation, Latent Dirichlet allocation for topic evaluation,
and grammar analysis for the readability evaluation.

\citet{deng2020two} used a Sequence-to-Sequence generation model and an adversarial model in the text summarization for Chinese texts.

\citet{zhang2019pretraining} used the Transformer network in encoder-decoder framework for the text summarization task, and
achieved a Rouge score of 33.48\% in the CNN/DailyMail dataset for news summarization.

\citet{xiao2019extractive} used the extractive LSTM-Minus model in the text summarization task, and by combining
global document level, sentence level and topic segment representation in the network, achieved a Rouge score of 31.99\%
in the Pubmed summarization dataset.

\citet{liakata2013discourse} used the CRF and SVM models for the sentence extractive summarization for the scientific articles.
\citet{cohan2015scientific} use the important sections in scientific papers as criteria for selecting important sentence.

\citet{sutskever2013importance} started to use sequence-to-sequence neural networks in abstractive summarization.
\citet{vinyals2015pointer} introduced copy mechanism from source to summary in sequence-to-sequence models.

\begin{figure*}
\begin{center}
  \fbox{\rule{0pt}{2in} \includegraphics[width=0.9\linewidth]{SummaryRule.pdf}}
\end{center}
   \caption{Rule based segment and importance based summarization.}
\label{fig:short}
\end{figure*}

\section{Hierachical Multi-grained Summarization}
\label{sec:hireachy}

Documents have their own organizations by different purposes, take the civil legal documents as an example,
There are 123 different kinds of arguing reasons in the Chinese Court Sentence Documents.
Largely list as follows:
\begin{itemize}
\item Disputes over personality rights,
\item Marriage, family, inheritance disputes,
\item Property rights disputes,
\item Contract, management without cause, disputes over unjust enrichment,
\item Labor disputes, personnel disputes,
\item Maritime disputes,
\item Civil disputes related to companies, securities, insurance, bills, etc.
\item Inciting disputes
\end{itemize}
However, each of these documents follow the same organization of segments, for example,
the purpose, the facts, the court opinions and the trail results.
Thus, for a meaningful text summarization includes all there segments of importance and ignore other segments.

\subsection{Segment Extraction}
The document of the court sentence usually states the arguing point of the two entities.
In this paper, a dictionary based arguing reason extractor was designed.
A total of 123 arguing reasons were classified by the legal experts out of 70 million
civil documents collected from the Chinese Court Sentence website.

Then I use several rule-based filters to extract meaningful segments. As Figure-1 shows, the document has been split into several
paragraphs, the paragraph with important words are kept and other paragraphs are discarded.

\subsection{Extractive Sentence Summarization}

For the sentence importance scoring model, the experts wrote the summarization for each legal document.
Base on the written summarization, each sentence importance can be easily computed by checking the overlap words.
I use the pre-trained Chinese language model, which has multiple layers of transformers,
for encoding sentences.
As Figure-2 shows, the model layers are organized as follows:
Take the last two layers of the RoBERTa encoding, then do the concatenation and take the Maximum element from the character
sequence as representation for each sentence, then concatenating to a dropout layer, a softmax layer,
and the output layer.

The learning rate is $10^{-4}$, the max sequence length is 50, the dropout rate is 5\%, and the batch size is 1.
The epoch number is 3. The experiment was carried on a Google Cloud TPU v3, with 32GB of RAM, and 8 chips with 16GB of the high speed of memory each,
which can provide 420 tera-flops of computation capability.

The classification dataset is unbalanced, thus I copy the positive samples multiple times to balance the dataset.
In total, there are 782,895 samples for training and 42 samples for test in the Legal Sentence Importance dataset.
The test accuracy is 92.06\%.

\subsection{Sequence To Sequence Summarization}

For the token generation in the summarization step, I use a sequence-to-sequence model.
By tagging each token existence in the corresponding summary, a sentence has its unique sequence of code for the token importance.
For the 14,908 samples of important sentences for training and 108 samples for test in the Legal Token Summarization Dataset, I design a token summarization model as follows:
First, encode the sentence by RoBERTa language model, and concatenate the first and last layers as the sentence representation.
Then concatenate to a CapsuleNet layer, then concatenates the output layer.

The learning rate is $10^{-3}$, and the bach size is 2, and the accumulation step is 8.
The training epoch is 10, and the accuracy of the best model in the test set is 42.05\%.
From Table(1) we can see that different number of transformer layers has no significant performance improvement for accuracy in the test set.
The RBT3, BRT3L and RoBERTa vary in number of neurons and number of layers,
which was popular pre-trained language models trained on 4.5 billion of Chinese tokens by \citet{cui2020revisiting}.

\begin{table}
\centering
\begin{tabular}{cccc}
\hline
\textbf{Model} & \textbf{Neurons}& \textbf{Layers} & \textbf{Accuracy} \\
\hline
RoBERTa & 768 & 12 & 42.05\% \\
RBT3 & 768 & 3 & 40.18\% \\
RBT3L & 1024 & 3 & 41.22\% \\
\hline
\end{tabular}
\caption{Different Pre-trained Language models accuracy in the sequence to sequece training steps.}
\end{table}

For balancing the readability and quality of the summary, I did not use the sequence predicted by the token summarization model directly.
Instead, a bigger threshold for the output of the model can improve the readability and the quality.
As a practical result, I choose 87.5\% as the threshold for keeping the important token.
That is to say, keeping the most important 87.5\% tokens for each inference shall get the best summarization result.

\begin{figure*}
\begin{center}
  \fbox{\rule{0pt}{2in} \includegraphics[width=0.9\linewidth]{SummaryModel.pdf}}
\end{center}
   \caption{Different architectures for multi-grained summarization.
   (a) Extractive Sentence Summarization,
   (b) Sequence To Sequence Summarization.}
\label{fig:short}
\end{figure*}

\subsection{Summarization Evaluation}

\citet{lin2004rouge} represents a set of similar metrics such as ROUGE-N, ROUGE-L, ROUGE-W, ROUGE-S, and ROUGE-SU.
In this paper, I use ROUGE-1, ROUGE-2, ROUGE-L and ROUGE-W as the summarization evaluation metrics.
The ROUGE-W score is computed using Formula (1).

\begin{equation}
  {R_W} = {R_1} * 0.2 + {R_2} * 0.4 + {R_L} * 0.4
\end{equation}

\begin{table}
\centering
\begin{tabular}{cccc}
\hline
\textbf{R-1} & \textbf{R-2} & \textbf{R-L} & \textbf{R-W} \\
\hline
32.62\% & 15.63\% & 31.79\% & 25.49\% \\
\hline
\end{tabular}
\caption{F-score for text summarization in the dataset.}
\end{table}

\section{Conclusion}

In this paper, I introduced a novel summarization method, which leveraging hierarchical features of the document.
By extracting multi-grained level from the document, the summary contains the paragraph-level, sentence-level and token-level key information.
The evaluation results showed that the multiple layers of transformers have improved the Rouge evaluation performance
and achieved a Rouge-W score of 25.49\% on the Chinese Legal Case Abstraction dataset.

\bibliography{anthology,acl2020,acl_summary}
\bibliographystyle{acl_natbib}

\end{document}