### 2021-01-16 A-BasicCommom/ACL2021/acl_word_segment.tex
1.acl2021 CN word segment.

%
% File acl2021.tex
%
%% Based on the style files for EMNLP 2020, which were
%% Based on the style files for ACL 2020, which were
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2021}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\renewcommand{\UrlFont}{\ttfamily\small}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Extra Large Sequence Transformer Model for Chinese Word Segment}

\author{Dezhou Shen \\
  Department of Computer Science\\
  Tsinghua University\\
  Beijing, CN 100084\\
  \texttt{sdz15@mails.tsinghua.edu.cn} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
 Chinese word segment is widely used in document analysis.
 In this paper, combining the Transformer-XL layer, the Fully-Connect layer, and the Conditional Random Field layer,
 the proposed model improved 5\% comparing to the BERT+CRF model in the MSR2005 Chinese word segment test dataset.
\end{abstract}

\section{Introduction}

 As a fundamental study for the semantic document segment, the Chinese word segment has been studied for decades.

 By tagging the status of each token, intended entities with longer than one character could be drawn out of documents,
 and unintended tokens are assigned special token type which are one character each.

\section{Related Work}

  \citet{devlin2019bert} proposed Bidirectional Encoder Representations from Transformers, and achieved 92.8\% on the CoNLL-2003 test set by
fine-tuning BERT on the downstream Named Entity Recognition task.

  \citet{yang2019xlnet} proposed an autoregressive language model which overcomes the limitations of BERT in sequence length,
results showed that XLNet achieved promising results on question answering, natural language inference, sentiment analysis, and document ranking tasks.

  \citet{tian2020improving} showed that extra trained BERT language model as encoder concatenate to the CRF can achieve 98.28\% to 98.40\% on the MSR2005 Chinese word segment test dataset.

  \citet{cui2019hierarchically} proposed a hierarchically-refined label attention network in Named Entity Recognition task, the result showed that LSTM+CRF is less competitive than their LSTM+LAN model on the WSJ,
Universal Dependencies v2.2, OntoNotes 5.0, and CCGBank dataset.

  \citet{Lafferty2001Conditional} proposed the Conditional Random Fields probabilistic model to segment and label sequence data, and outperforms the hidden Markov models and maximum entropy Markov models by 0.14\%~0.82\% in accuracy on the Penn treebank POS dataset.

\section{Methodology}
\label{sec:hireachy}

\subsection{Data Augment}

For the Chinese 8-bit Unicode Transformation Format encodings, there are two different encoding for the same printable ASCII tokens, one is named as the half character encoding,
and the other is named as the full character encoding.
Their difference relies on that one token take one byte or three bytes. The MSR2005 Chinese word segment dataset used the full character encoding, but in most cases,
people tend to use the half character encoding to comply with ASCII code for compatibility reason.
In this paper, I use a conversion function to convert the MSR2005 training dataset tokens to half character encoding for the tokens which have the corresponding tokens in the ASCII table.
Thanks to the data augment method, training dataset sentences size is increased from 86,924 to 166,740.

\subsection{Tokenization, Encoding And Labeling}

For a document with multiple sentences and paragraphs, the first step is splitting the document into lines by carefully choosing the separator token.
For Chinese sentences, the line-breaks, commas, periods are the most common separator tokens.
Before encoding the tokens, sentences are padded with padding token [0] to the max sequence length, I choose 128 as the max sequence length for the MSR2005 dataset.
After tokenization, I use the XLNet encoder to represent each token as the token integers.
Each token has been labeled according to its type, for the one character entity, its status is labeled to [0].
For the longer than one character entity, the first character is labeled to [1], the last character is labeled to [3], and the rest characters are labeled to [2].

\subsection{Sequence Encoder}

After the encoding steps, the tokens are encoded to token integers which represents their vocabulary indices.
To encode by XLNet, the mask tokens and the segment tokens are prepared.
The non-padding tokens' mask tokens are set to [1], and the segment tokens are set to [0].
The padding tokens' make tokens are set to [0], and the segment tokens are set to [1] for telling the model the boundary of the sentences.
For encoding the sequence, all the layers are preserved for later usage.

\subsection{Sequence Hidden Representation}

For a naive solution, it is perfect to add a softmax layer or a CRF layer.
However, I designed a multiple layers to take advantage of the different layers for the transfer learning.
Firstly, take the first layer, and the last layer of the XLNet sequence encoding.
Then use two dropout layers with each dropout rate pf 5\%, which means the probability of an element to be zeroed.
Thirdly, I use a fully connected layer to map the hidden size of XLNet to the status types, which is 4 in the Chinese word segment task.
Lastly, take the logarithm, and a softmax conversion before concatenating to the CRF layer.

\subsection{Conditional Random Field}

Conditional Random Field is a traditional method to keep the output of the sequence following a reasonable pattern.
Each token sequence has a corresponding mask sequence to show its margin, the token [True] shows that the token is a non-padding token,
and the token [False] shows that the token is a padding token.

\subsection{Additional Loss function}

Besides, the Conditional Random Field loss for training, I added a type loss for training speed.
After the two dropout layer, an extra fully connected layer was concatenated to the last dropout layer to get the type loss.
Then use a multi-label one-versus-all loss based on max-entropy to get the sequence type loss.
The type sequence target was smoothed by 0 and 1 values, the type [0] for the token status of [0], and the type [1] for the other token status.

\begin{figure*}
\begin{center}
  \fbox{\rule{0pt}{2in} \includegraphics[width=0.9\linewidth]{SummaryModel.pdf}}
\end{center}
   \caption{Different architectures for multi-grained summarization.
   (a) Extractive Sentence Summarization,
   (b) Sequence To Sequence Summarization.}
\label{fig:short}
\end{figure*}

\section{Experiments}

\subsection{Common Setups}

\par The experiment was performed on a Google Cloud NVIDIA V100 with 16GB of the high speed of memory each, with 32GB of RAM,
which can provide 112 tera-flops of computation capability.

The learning rate is $4^{-5}$, and the batch size is 64, for the BERT and XLNet.
The learning rate is $2^{-3}$, and the batch size is 512 for the bidirectional LSTM.
The max sequence length is 128, and the training epoch is 10.
The RoBERTa in this paper is popular pre-trained RoBERTa language models trained on 4.5 billion of Chinese tokens by \citet{cui2020revisiting}.
The XLNet in this paper is another pre-trained XLNet language models trained on 4.5 billion of Chinese tokens by \citet{cui2020revisiting}.

\begin{table}
\centering
\begin{tabular}{cccc}
\hline
\textbf{Model} & \textbf{Neurons}& \textbf{Layers} & \textbf{macro-F1} \\
\hline
RNN & 256 & 2 & 92.62\% \\
RoBERTa & 768 & 12 & 92.62\% \\
XLNet & 768 & 12 & 97.88\% \\
\hline
\end{tabular}
\caption{Pre-trained language models macro-F1 scores for the Chinese word segement task.}
\end{table}

All the Table (1) models share almost the same network architecture except for the embedding of the tokens.

\section{Conclusion}

In this paper, I introduced a novel Chinese word segment architecture, which leveraging hierarchical layers of the document encoding layers.
By keeping the Conditional Random Field layer, I proposed a novel sequence hidden representation of the document sequence, which achieved 5\% performance gain in macro F1-score than
the bidirectional lstm and CRF architecture on the MSR2005 Chinese word segment test dataset.

\bibliographystyle{acl_natbib}
\bibliography{anthology,acl2021,acl_word}

%\appendix

\end{document}